Light and growth form interact to shape stomatal ratio among British angiosperms

### 2017-10-21 ms/ms.Rnw
revised introduction

% MARKER (still revising, adding predictions here): Specifically, I predicted that species with faster life histories, especially therophytes (annuals), would have greater \els~than species with slower life histories, especially phanerophytes (shrubs and trees).

The final question is important for addressing whether amphistomy is part of a coordinated syndrome of traits that promote higher photosynthetic rate, as both the light and growth form hypotheses assume. If evolved increases in stomatal ratio are mediated by shifting abaxial stomata to the adaxial surface, holding total stomatal density constant, then the overall increase in CO$_2$ diffusion would be small. In contrast, if amphistomy evolves by increasing adaxial stomatal density while holding abaxial density constant, then \textit{total} stomatal density must increase as well. Evolutionary coordination of amphistomy and high stomatal density would reinforce one another, increasing CO$_2$ supply to chloroplasts more than changes in either trait would in isolation. Understanding selection on coordinated traits can explain the evolution of major functional trait axes and syndromes.

>>>>

In answering these questions, I both reassessed previous hypotheses using newer phylogenetic comparative methods and evaluated previously untested hypotheses. I predicted \textit{a priori} that light habitat and growth would be correlated. Species with faster life histories, especially therophytes (annuals), would on average inhabitat sunnier environments than species with slower life histories, especially phanerophytes (shrubs and trees). Based on hypotheses from previous studies, I also predicted that herbaceous growth form and high light would be associated with amphistomy, even after controlling for phylogenetic nonindependence. Although these predictions have been tested previously, it is critical to reevaluate them here with updated methods because the subsequent untested hypotheses build on these results. The first novel hypothesis I tested predicts that light and growth form interact. Specifically, I hypothesized that both high light and herbaceous growth would be required to favor a more even stomatal ratio (i.e. amphistomy). Finally, I tested whether amphistomy is part of a coordinated syndrome of traits that promote higher photosynthetic rate. If high light and growth form favor amphistomy because it increases photosynthesis, then it follows that they should also favor other stomatal traits that reinforce this advantage. If evolved increases in stomatal ratio are mediated by shifting abaxial stomata to the adaxial surface, holding total stomatal density constant, then the overall increase in CO$_2$ diffusion would be small. In contrast, if amphistomy evolves by increasing adaxial stomatal density while holding abaxial density constant, then \textit{total} stomatal density must increase as well. Evolutionary coordination of amphistomy and high stomatal density would thus reinforce one another, increasing CO$_2$ supply to chloroplasts more than changes in either trait would in isolation. Understanding selection on coordinated traits can explain the evolution of major functional trait axes and syndromes.
